Exercise 2: Quality Attributes and Architectural Design of the HS07 System

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Section 5.1 and 5.2 has been merged into 4.1 and 4.3 respectively. Needs more updating.

\subsection{Refinement}
The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.

@@

% =====================================================================
\section{Evaluation}
The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
is discussed in the following chapters.
\subsection{Refined Scenario \#1}
The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.

\subsubsection{Adding new sensor/actuator}
Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.

\subsubsection{Adding different types}
Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.

We have chosen to focus in the integration of new types of thermometers and actuators.

@@

@@

\subsection{Refined Scenario \#1 - Tactics}
\subsubsection{Modifiablity}
\begin{itemize}
\item New Thermometer/Radiator implementation
\begin{itemize}
\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification. Changing the ThemometerService to use e.g. AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService. At present the getController returns an object and the AbstractFactory may be implemented to also return this. It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method). This imply several tactics, as it \emph{localize modifications} to the temperature implementation. If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future. \emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class. 
\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has. There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery). It may be updated with a DNS-lookup to make this weaker.
\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
\end{itemize}
\item New sensor/actuator type
\begin{itemize}
\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types. The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
\end{itemize}
\end{itemize}
\subsubsection{Testability}
Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing. The test harness for the HS07 system will consist of a series of JUnit test cases on a modular level. On a larger scale, the deployment on a PC, shown in /cite{SAiP1H1} is one big test scenario where sensors and actuators a stubbed using visual inspection for verification. This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
\subsection{Refined Scenario \#3 - Tactics}
\subsubsection{Availability}
As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g. \emph{voting}). At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}. Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen. As an example of a design may be mentioned the Apache web-server design, which is redundant processes. A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive. If this happens it is simply killed and an identical process is spawned in its place.

>>>>

\subsection{Refinement}
\label{sec:refinement}

The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.

@@

% =====================================================================
\section{Evaluation}

The following sections will elaborate on the refined scenarios described in section \ref{sec:refinement}.
The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined
scenarios will be described and tactics that can be applied to the architectural design in order to
fulfill the quality attributes are identified.

\subsection{Refined Scenario \#1}
The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.

\begin{description}
\item[Adding new sensor/actuator:] Localized to changes within the given sensor or actuator type.
Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case. 
\item[Adding different types:] Update by addition will need to be built into the architecture in the sensor
and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and
thermometers are hardcoded into the gateway, it will need to be more abstract.
\end{description}

We have chosen to focus in the integration of new types of thermometers and actuators and
addition of different sensor and actuator types will not be described any further.

\subsubsection{Modifiablity}
\begin{itemize}
\item New Thermometer/Radiator implementation
\begin{itemize}
\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification. Changing the ThemometerService to use e.g. AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService. At present the getController returns an object and the AbstractFactory may be implemented to also return this. It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method). This imply several tactics, as it \emph{localize modifications} to the temperature implementation. If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future. \emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class. 
\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has. There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery). It may be updated with a DNS-lookup to make this weaker.
\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
\end{itemize}
\item New sensor/actuator type
\begin{itemize}
\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types. The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
\end{itemize}
\end{itemize}
\subsubsection{Testability}
Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing. The test harness for the HS07 system will consist of a series of JUnit test cases on a modular level. On a larger scale, the deployment on a PC, shown in /cite{SAiP1H1} is one big test scenario where sensors and actuators a stubbed using visual inspection for verification. This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.

@@

\subsubsection{Availability}
As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g. \emph{voting}). At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}. Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen. As an example of a design may be mentioned the Apache web-server design, which is redundant processes. A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive. If this happens it is simply killed and an identical process is spawned in its place.

@@
